---
title: "Latent Variable Measurement Error Correction with `lpmec`"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lpmec Package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[UTF-8]{inputenc}
---

## Introduction

This tutorial demonstrates how to use the `lpmec` package for measurement error correction in regression models using latent variable estimation. The package implements bootstrapped analyses to account for measurement error in observed indicators and provides corrected regression coefficients.

## Installation

First install the required dependencies and the `lpmec` package:
```{r setup, eval=FALSE}
# Install lpmec from source (replace with appropriate installation method)
# devtools::install_github("cjerzak/lpmec-software", subdir = "lpmec")
```


## Basic Usage

### Data Simulation

Simulate data with a latent predictor and observed binary indicators:

```{r, eval=TRUE}
set.seed(123)
n <- 1000  # Number of observations
d <- 10    # Number of observable indicators

# Generate latent variable and observed outcomes
x_true <- rnorm(n)
Yobs <- 0.4 * x_true + rnorm(n, sd = 0.35)

# Generate binary indicators of latent variable
ObservablesMat <- sapply(1:d, function(j) {
  p <- pnorm(0.5 * x_true + rnorm(n, sd = 0.5))
  rbinom(n, 1, p)
})
```

### Running the Analysis

Use lpmec to estimate corrected coefficients:

```{r, eval=TRUE}
library(lpmec)

# Run analysis with bootstrap uncertainty estimates
results <- lpmec(
  Y = Yobs,
  observables = as.data.frame(ObservablesMat),
  n_boot = 10,      # Use 0 to disable bootstrap resampling
  n_partition = 5,  # Reduced for demonstration
  estimation_method = "em"
)
```

Compare naive and corrected estimates:
```{r, eval=TRUE}
print(results)
summary(results)
```

You can visualize the relationship between split-half estimates:
```{r}
plot(results)
```

### Inference for Median Partition Aggregation

The default partition aggregation is the finite-partition median. Because the
median is nonsmooth, the formal route for uncertainty is subsampling or
m-out-of-n bootstrap with root-scaled intervals:

```{r, eval=FALSE}
results_med <- lpmec(
  Y = Yobs,
  observables = as.data.frame(ObservablesMat),
  n_boot = 499L,
  n_partition = 10L,
  estimation_method = "em",
  partition_aggregation = "median",
  bootstrap_method = "subsampling",
  boot_m_rule = "power",
  boot_m_exponent = 0.70,
  boot_ci_type = "root",
  seed = 123
)
```

This path holds the same realized partition set fixed across the original
sample and all subsamples, reruns the full latent-score and correction pipeline
inside each subsample, and rescales the empirical root statistic from `m` to
the original `n` rate. Inspect `boot_m`, `boot_m_ratio`,
`valid_partitions_by_boot`, `bootstrap_success_rate`, and sensitivity across a
grid of `boot_m` values in applied work.

## Advanced Features

### Using Different Estimation Methods

The package supports multiple estimation approaches:

```{r, eval=TRUE}
# Bayesian MCMC estimation with the default pscl backend
if(FALSE){
mcmc_results <- lpmec(
  Y = Yobs,
  observables = as.data.frame(ObservablesMat),
  estimation_method = "mcmc"
)

# To use the optional NumPyro backend, first run build_backend()
# or provide your own Python environment with JAX and NumPyro.
numpyro_results <- lpmec(
  Y = Yobs,
  observables = as.data.frame(ObservablesMat),
  estimation_method = "mcmc",
  mcmc_control = list(backend = "numpyro"),
  conda_env = "lpmec"
)
}
```

## Conclusion

The `lpmec` package provides robust measurement error correction through:

  - Bootstrapped latent variable estimation
  - Multiple correction methods (OLS, IV, Bayesian)
  - Flexible estimation backends (EM, MCMC)

Refer to package documentation for advanced configuration options.
